What can Stack Traces tell us about Java Exceptions?

### 2014-08-13 android-stacks.tex
including more detailed examples of strange wrappings.

\begin{document}
\title{What Can Stack Traces Tell us about of the use Exceptions in Java?}

@@

\subsection{Georgios's and Roberta's Merged Introduction}

@@

\subsection{Error Handling in Programming Languages}

Techniques for error detection and handling are not  an optional add-on but a
fundamental part of any system~\cite{bruntink2006discovering}. Several
approaches exist for detecting errors and responding to them, such as the return-code 
idiom (widely used in systems written in C), and the exception handling 
mechanism, embedded in many main-stream programming languages.
The main disadvantage of return codes is the fact that the programmer can just
ignore the return code and let the program execute in inconsistent state. 
The exception mechanism~\cite{goodenough1975exception} was proposed as a way of
preventing errors from being ignored and was embedded in many mainstream
programming languages ~\cite{garcia2001comparative} since then.

@@

Both Java [ref: Gosling et al. 1996] and CLU [Liskov and Snyder 1979] support the checked exceptions,
 hence the compiler can then check if handlers are provided to such exceptions.
 However, this support is only partial because both languages also provides a form of unchecked exceptions.
In CLU,  if a handler is found for a given exception, it is converted in a predened default exception, 
called failure which is the only one which can be implicitly propagated [ref-clue].
 Java allows a hybrid approach on which the developer can define specific unchecked and checked exceptions,
and use both kinds of exceptions inside a single system.
In both languages the developer of a client module is not warned by the compiler if 
 unchecked exceptions are thrown by a module.

[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems

Although exception handling mechanism has been embedded in several programming languages 
as show above,   the exception handling code is often generally poorly understood and the least
tested part of software systems.

There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
are related to this debate}. The debates, however, are based on the developers own experience in
using both models. There is a lack of empirical study tackling the pros and cons of each model.

@@

In Java, an exception is represented according to the class hierarchy shown in
Figure~\ref{fig:exchier}.  According to it every exception is an
instance of the Throwable class, and can be of three kinds: checked exceptions
(extends Exception), run-time exceptions (extends RuntimeException), and errors
(extends Error)~\cite{gosling2000java}.

@@

Neither run-time exceptions nor errors need to be specified on method
signatures, and hence both are referred as unchecked exceptions. The Java
Virtual Machine represents as runtime exceptions invalid operations detected in
the program (out-of-bounds array index, divide-by-zero error, null pointer
references) from which most of the programs are not expected to recover. By
convention an Error represents an unrecoverable condition which usually results
from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
normally should not be handled inside the application.

The Java specification~\cite{gosling2000java} suggests that user-defined
exceptions should be checked exceptions. The main reason is because doing so the
callers of a method will know about the exceptions that a method can throw, and
so they can decide what to do about them.

In Java an exception can be thrown in one of the following
circumstances~\cite{gosling2000java}: (i) explicitly thrown when a throw statement is executed; 
(ii) implicitly thrown by the JVM when the evaluation of an expression
 violates the normal semantics of language (e.g., out-of-bounds array index,
division-by-zero, access to a null reference); or (iii) implicitly thrown by
the JVM due to an internal error or resource limitation (e.g.,
OutofMemoryError).

@@

When an exception is throw it causes a transfer of control from the point
where the exception occurred to a point that can be specified by the programmer
(exception handler) - in Java it is represented by the try-catch block .

A common way of  propagating exceptions in Java programs, is the exception wrapping (also known as exception chaining), which allows one exception to be wrapped in another exception and re-thrown. 
Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
exception wrapping. The bottom part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

\subsection{Exception Handling Best Practices in Java}
\label{sec:best}

Several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}. 
A list of the guidelines related to exception types and exception
propagation in Java the following:\footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}

@@

The main goal of this study was to gain a better understanding
of the use of exceptions in Java programs based mainly on the information
available on exception stack traces embedded on Github issues.
The exceptions available on exception stack trace are not limited to the exceptions 
expliciltly thrown by the Java developer but also include
the exceptions implicitly thrown by the JVM during the program execution,
due to a programming error (e.g., out-of-bounds array index, division-by-zero, 
access to a null reference) or an internal error or resource limitation (e.g.,
OutofMemoryError).

@@

  \item[RQ1] What are the common characteristics of such stack traces?

  \item[RQ2] What patterns of exception (mis)-use emerge from the stack trace
    analysis?

@@

@@

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of
projects analyzed we opted for not restricting the analysis only on defect
issues.\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{www.dimap.ufrn.br/~roberta/icsme2014}}

@@

After distilling the information available on the 28,800 stack traces were extracted from the 
issues and issues comments of 16,836 Java projects: 607 distinct root causes were found, 
and  2,026 distinct top level exceptions were found.

From this set the java.lang.NullPointerException  was the root cause most 
reported;  7578 stack traces had a java.lang.NullPointerException  as root cause,
 which corresponds of 26,31\%  of all analyzed stack traces.

If we perform a cross-project analysis, the java.lang.NullPointerException, we could observe that such 
exception was reported on 1,671 projects (which represents 44,47\% of all projects reporting at least one stack trace).

@@

To better understand the root causes reported on stacks, we performed a categorization 
study on which the top 100 root causes reported on stack-issues were manually inspected.
The characterization process consisted of classifiying each exception in one of the
chategories listed on Table X, based on the inspection of source code, Javadoc, 
or additional documentation related to the exception.

\begin{table}
  \centering
  \begin{tabular}{|p{2cm}| p{5cm}|}
    \hline
    \bfseries{Category} & \bfseries{Description} \\
    \hline
      Programming logic &  Exceptions implicitly  thrown by the JVM when the 
evaluation of an expression violates the normal semantics of language (e.g., 
out-of-bounds array index, division-by-zero, access to a null reference). 
Such exceptions are usually defined on java.lang and java.util packages. \\ \hline
      Resources (IO)                         & Exception releated to the manipulation of IO resources or device  (e.g., file, network) \\ \hline
      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
      Reflection                             & Exceptions related to Java reflection library  \\ \hline
      Specific                & Exceptions created for specific applications, frameworks or libraries \\ \hline
      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
    \hline
  \end{tabular}
  \caption{Categories description.}
  \label{tab:categories}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{llll}
    \hline
    \bfseries{Root Cause} & \bfseries{Wrapper Type} & \bfseries{Occurrences} & \bfseries{Projects} \\
    \hline
     Runtime    & -   &  2360 & 359 \\

    Runtime   & Runtime & 560  & 182 \\

 Checked    &   -  & 422  & 117 \\
     Error    &   -   & 381  & 141 \\
   Checked &  Runtime   & 109  & 70  \\
    Checked   & Checked & 98   & 42  \\
   Error   &  Runtime   & 55   & 38  \\
   Runtime &  Checked   & 22   & 12  \\
    Error     & Error   & 15   & 14  \\
  Undefined    &  -  &  15   & 10  \\
   Runtime & Error     &  13   & 6 \\
  Error   & Checked   &  8    & 7 \\
    Checked &Error     &  7    & 4 \\
    Runtime &Throwable &  3    & 1 \\
   Runtime & Undefined &  3    & 3 \\
   Throwable    & -  &  1    & 1 \\
   Error   & Undefined &  1    & 1 \\
    \hline
  \end{tabular}
\caption{Kinds of wrappings found on the stacks of 482 ANDROID projects found on Github}
\label{tab:wrappingandroid}
\end{table}

@@

@@

Since there is no way of enforcing Java exception type conventions during program development,
we could observe the stack trace analysis combined with types these three types
of exceptions were used interchangeably. This interchangeability makes the
behaviour of the exception handling code more complex and less reliable. For
instance, considering a situation were an instance of OutOfMemoryError (a situation that
should not be handled) was wrapped in a checked exception (a situation that must
he handled), in this case this wrapping is telling to the caller that it should
handled.

@@

The exception (mis-)use patterns that emerged  in this study sheds light on 
problems that may be reducing the robustness of current Java programs in the face of
exceptional conditions. \noindent\emph{The undocumented runtime exceptions} thrown by library code -
documenting runtime exceptions is a tedious and error prone task, to help developers
mitigate this problem tools should be developed to automate the documentation of runtime exceptions
scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}. 
 \noindent\emph{The odd wrappings}  found in this study can lead to what we call
an \emph{exception handling confusion problem} which can lead the program to
an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
 one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the JVM which make impossible the program to recover. Further investigation is needed on finding 
ways to help developers in deadling with such problem, either preventing odd wrappings or enabling 
the developer to better dealing with them. Last but not least, the high number of null pointer exceptions 
lead us to think on other pertinent research direction: whether language designs which avoid null pointers, 
such as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs.

TO DISCUSS: How can we benefit from such cross-project crash analysis?

@@

@@

\section {Old Introduction}

....... OLD VERSION ......

>>>>

\begin{document}
\title{ What can Stack Traces tell us about Java Exceptions? }

%\title{A first Glance of what is causing Mobile Crashes: A GitHub and Googlecode Repository Mining Study }

@@

@@

\subsection{Exception Handling in Programming Languages}

Techniques for error detection and handling are not  an optional add-on but a
fundamental part of any system~\cite{bruntink2006discovering}. 
The exception mechanism~\cite{goodenough1975exception} was proposed as a way of
preventing errors from being ignored and was embedded in many mainstream
programming languages ~\cite{garcia2001comparative} since then.

@@

Java [ref: Gosling et al. 1996] and CLU [Liskov and Snyder 1979] support the checked exceptions,
 (therefore the compiler can then check if handlers are provided to such exceptions), however, this support is 
only partial because both languages also provides a form of unchecked exceptions [ref: Robilard].
In CLU,  if a handler is not found for a given exception, it is automaticaly converted in a predened default exception, 
called failure which is the only one which can be implicitly propagated [ref-clue].
 In Java the developer can define specific unchecked and checked exceptions and use both kinds in a single system.

There is a long-lasting debate about the pros and cons of checked and unchecked exceptions 
~\cite{javatut,stackoverlow,debate},\footnote{152 questions in Stackoverlow
are related to this debate}. However, the debates are mainly based on the developers own experience in
using both models, since there is still few empirical evidence of the pros and cons of each model.

%Although exception handling mechanism has been embedded in several programming languages 
%as show above,  the exception handling code is often poorly understood and the least
%tested part of software systems [ref:POORLY].

%[REF]: Static Analysis to Support the Evolution of Exception Structure in Object-Oriented Systems

@@

In Java, an exception is represented according to the class hierarchy shown in
Figure~\ref{fig:exchier}.  According to it every exception is an
instance of the Throwable class, and can be of three kinds: checked exceptions
(extends Exception), run-time exceptions (extends RuntimeException), and errors
(extends Error)~\cite{gosling2000java}. Neither run-time exceptions nor errors need to be specified on method
signatures, and hence both are referred as unchecked exceptions.

@@

The Java Virtual Machine represents as runtime exceptions invalid operations detected in
the program (out-of-bounds array index, divide-by-zero error, null pointer
references) from which most of the programs are not expected to recover. By
convention an Error represents an unrecoverable condition which usually results
from failures detected by the Java Virtual Machine, such as OutOfMemoryError and
normally should not be handled inside the application. The Java specification~\cite{gosling2000java} 
suggests that user-defined exceptions should be checked exceptions. The main reason is because doing so the
callers of a method will know about the exceptions that a method can throw, and
so they can decide what to do about them.

Hence, in Java an exception can be thrown in one of the following
circumstances~\cite{gosling2000java}: 
(i) explicitly thrown when a throw statement is executed; 
(ii) implicitly thrown by the JVM when the evaluation of an expression
 violates the normal semantics of language, also refered as a coding error
 (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
OutofMemoryError).

@@

When an exception is throw, it causes a transfer of control from the point
where the exception occurred to a point that can be specified by the programmer
(exception handler) - in Java it is represented by the try-catch block . A common way of  
propagating exceptions in Java programs is through exception wrapping
 (also known as exception chaining). Exception wrapping allows one exception to be wrapped in another one and re-thrown. 
Figure~\ref{fig:wrapping} presents a stack trace associated which illustrates an
exception wrapping. The bottom part of the stack trace is the \emph{root cause}, which indicates the
first reason for the error to be thrown (in this case, the computer run out of
memory). The top part of the stack trace indicates the location of the exception
manifestation (which will call \emph{top level exception} along this paper). The
execution flow  between the root cause and the exception manifestation may
include other intermediate exception wrappings. In all levels, the exception
\emph{signaler}, is the method that threw the exception, represented on the
stack trace as the first method call below the exception declaration.

@@

@@

Based mainly on the information available on exception stack traces embedded 
on Github issues, the goal of this study was to gain a better understanding 
of exception-related system crashes.

The exceptions available on exception stack traces are not limited to the exceptions 
expliciltly thrown by the Java developer but also include
the exceptions implicitly thrown by the JVM due to programming errors (e.g., out-of-bounds array index, division-by-zero, 
access to a null reference) or resource limitation (e.g., OutOfMemoryError).

@@

  \item[RQ1] What are the common characteristics of exception stack traces reported on issues?/What are the main causes of crashes reported on issues? (more explicit?)

  \item[RQ2] What patterns of exception (mis)-use emerge from the exception stack trace
    analysis?

@@

\subsection{Meaning of Exception Types and related Best Practices}
\label{sec:best}

Several general guidelines have been proposed on how to use Java
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}. 
We compiled a list of the guidelines related to exception types and exception
propagation in Java the following:\footnote{We also compiled guidelines related
to exception handling but they are out of the scope of this paper.}

@@

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed, we opted for not restricting the analysis only on defect
issues.\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{www.dimap.ufrn.br/~roberta/icsme2014}}

@@

After distilling the information available on the 28,800 stack traces (extracted from the 
issues and issues comments of 16,836 Java projects) we found 607 distinct exceptions
reported as the root causes of stack traces.

Table \ref{tab:toptenjava} illustrates the top 10 root couses found in the study ranked by the number of distinct
project repositories on which they were reported. The table illustrates the
number of times each exception was reported as the root cause of a stack trace
, the number of distinct projects reporting such causes on stack traces, and the percentages of
such in relation to all reported exceptions and analyzed projects respectaly.

\begin{table}
\begin{tabular}{lrrrr}
    \hline
    \bfseries{Exception} &  \multicolumn{2}{c}{\bfseries{Occurrences}} &  \multicolumn{2}{c}{\bfseries{Projects}}\\
    & \bfseries{\#} &  \bfseries{\%} & \bfseries{\# } & \bfseries{\% } \\
    \hline
java.lang.NullPointerException             & 7578 & 26,31\% & 1671 & 44,47 \\
java.lang.IllegalArgumentException         & 1573 & 5,46\%  & 670  & 17,83\% \\
java.lang.ClassNotFoundException           & 1203 & 4,18\%  & 514  & 13,68\% \\
java.lang.RuntimeException                 & 1093 & 3,80\%  & 411  & 10,94\% \\
java.lang.IllegalStateException            & 964  & 3,35\%  & 431  & 11,47\% \\
java.lang.ClassCastException               & 886  & 3,08\%  & 408  & 10,86\% \\
java.lang.ArrayIndexOutOfBoundsException   & 834  & 2,90\%  & 345  & 9,18\% \\
java.lang.NoSuchMethodError                & 822  & 2,85\%  & 369  & 9,82\% \\
java.io.IOException                        & 656  & 2,28\%  & 337  & 8,97\% \\
java.lang.OutOfMemoryError                 & 601  & 2,09\%  & 235  & 6,25\% \\
%java.lang.NoClassDefFoundError            & 554  & 1,92\%  & 299  & 7,96\% \\
%java.util.ConcurrentModificationException & 465  & 1,61\%  & 180  & 4,79\% \\
%java.lang.IndexOutOfBoundsException       & 445  & 1,55\%  & 205  & 5,46\% \\
\hline
All	 &	23956 &	100\%	& 3758 & 100\%	 \\
\hline
  \end{tabular}
\caption{Top 10 most popular exceptions on issues in Java projects}
\label{tab:toptenjava}
\end{table}

We can observe that most of this set is composed by exceptions defined by java.lang,
which are usually implicitly thrown by JMV dur to a programming mistake. From thi set
the java.lang.NullPointerException  was the most reported root cause;  7578 stack traces 
had a java.lang.NullPointerException  as root cause, which corresponds of 26,31\%  of 
all analyzed stack traces. Considering the number of distinct project repositories, 
the java.lang.NullPointerException, was reported on at least one issue of 1,671 projects 
(which represents 44,47\% of all projects reporting at least one stack trace).

@@

In Section XX we provide further discussion on the NullPointerException problem, pointing 
to ways of how to deal with it and mitigate it.

\subsubsection{Classifying Top 100 Exceptions}

To get a broather view of the root causes of stack traces, we performed an empirical evaluation
whose goal was to categorized the top 100 root causes.

After a first inspection of the complete list of exceptions reported, a first set of exception 
categories was defined. This list represent the most common terms reffered on exception
names and javadoc documentation. A similar approach was addopted in [COELHO-ECOOP].
The list of categories is illustrated in Table~\ref{tab:categories}. Hence in the second phase
of the study, each exception was mapped to one of the categories previously defined - 
this analysis was based on the inspection of source code, Javadoc, or additional 
documentation related to the exception.

To ensure the quality of the classification, three independent coders classified a randomly selected
sample of 25 exception types (from the total 100) using the same categories;
the interater agreement was high (96\%). We can therefore assume that the
categories are descriptive enough for covering the whole range of exception
types that can occur in real world Java projects.

\begin{table}
  \centering
  \begin{tabular}{|p{2cm}| p{5cm}|}
    \hline
    \bfseries{Category} & \bfseries{Description} \\
    \hline
      Programming logic &  Exceptions implicitly  thrown by the JVM when the 
evaluation of an expression violates the normal semantics of language (e.g., 
out-of-bounds array index, division-by-zero, access to a null reference). 
Such exceptions are usually defined on java.lang and java.util packages. \\ \hline
      Resources (IO)                         & Exception releated to the manipulation of IO resources or device  (e.g., file, network) \\ \hline
      Security                               & Exception releated to security issues (e.g. password validation, criptography) \\ \hline
      Concurrency                            & Exceptions related to multi-threaded programming \\ \hline
      Backward compatibility                 &  A list of exceptions related to backward compatibility - a complete list can be found here [ref]  \\ \hline
      Reflection                             & Exceptions related to Java reflection library  \\ \hline
      Specific                & Exceptions created for specific applications, frameworks or libraries \\ \hline
      General (Exception, Error, Runtime)    & The top level exceptions in Java    \\ \hline
    \hline
  \end{tabular}
  \caption{Description of categories.}
  \label{tab:categories}
\end{table}

@@

\begin{table}
\centering
\begin{tabular}{llll}
    \hline
    \bfseries{Wrapper Type} & \bfseries{Root Cause} & \bfseries{Occurrences} & \bfseries{Projects} \\
    \hline
      none   & Runtime   &  2360 & 359 \\
      none  &   Checked  & 422  & 117 \\
      none  &   Error   & 381  & 141 \\
      none &  Undefined  &  15   & 10  \\
      none & Throwable  &  1    & 1 \\
     \hline
      Runtime   & Runtime & 560  & 182 \\    
      Checked   & Checked & 98   & 42  \\
      Error     & Error   & 15   & 14  \\
    \hline
      Runtime &  Checked   & 114  & 70  \\
      Runtime   &  Error   & 57   & 38  \\
      Checked &  Runtime   & 22   & 12  \\
      Checked & Error    &  8    & 7 \\
      Error & Checked     &  7    & 4 \\
      Error & Runtime     &  13   & 6 \\
      Throwable & Runtime &  3    & 1 \\
      Undefined & Runtime &  3    & 3 \\
      Undefined & Error &  1    & 1 \\
    \hline
  \end{tabular}
\caption{Kinds of wrappings found on the stacks of 482 ANDROID projects found on Github}
\label{tab:wrappingandroid}
\end{table}

@@

\subsubsection{Runtime wrapping Checked Exception}

Such wrapping was detected on distinct 70 projects. Inspecting the elements responsible for performing such wrapping we could observe that,
most of such wrappings were perform on methods defined on Android.

ANDROID	 \- 66
APP	\- 29 	(HOW MANY?? AND WHICH??)
LIB  \- 19 (HOW MANY AND WHICH?)

Taking a closer look on the signalers reposponsible for wrapping we could observe that, aproximately 41%
of all wrappings were performed by a single class, called ActivityThread, responsible for managing the life cycle 
of an Android application. It performs the application callbacks.

android.app.ActivityThread \- 47
android.view.LayoutInflater \- 7
org.gege.caldavsyncadapter.syncadapter.SyncAdapter \- 7
org.springframework.web.client.RestTemplate \- 6
android.os.AsyncTask \- 5

Code snipet:

android.app.ActivityThread.

public final Activity startActivityNow(...) \{
...
return performLaunchActivity(r, null);
\}

private Activity performLaunchActivity(...) \{
try \{
.....
\} catch (Exception e) {
..
throw new RuntimeException(
"Unable to instantiate activity " + component
+ ": " + e.toString(), e);
..
\}
\}

The exceptions most wrapped on RuntimeExceptinos were:

(java.lang.RuntimeException,java.lang.ClassNotFoundException)	32
(java.lang.RuntimeException,java.lang.NoSuchMethodException)	6
(java.lang.RuntimeException,java.lang.InstantiationException)	5
(com.github.andlyticsproject.console.NetworkException,org.apache.http.client.HttpResponseException)	5
(android.view.InflateException,java.lang.NoSuchMethodException)	4

java.lang.RuntimeException	69 \- 61%
android.view.InflateException	10 \- 8,7%
com.github.andlyticsproject.console.NetworkException	7
java.lang.IllegalStateException	6
org.springframework.web.client.ResourceAccessException	6
com.github.andlyticsproject.console.DevConsoleProtocolException	3

\subsubsection{Runtime Exception wrapping Error}

 We could find  57 stack traces on 38 projects on which an Error was wrapped in a Runtime Exception.

ANDROID	50
LIB 	6
APP	1

android.os.AsyncTask \-	25
android.app.ActivityThread \- 15
android.view.LayoutInflater \- 7
org.asynctasktex.AsyncTaskExd \- 3
android.support.v4.content.ModernAsyncTask \- 2

EXAMPLE: android.os.AsyncTask

@Override
protected void done() \{
...
try \{
...

\} catch (InterruptedException e) \{
android.util.Log.w(.., e);
\} catch (ExecutionException e) \{
throw new RuntimeException("An error occured while executing doInBackground()",
e.getCause());
\} catch (CancellationException e)  \{
...
\} catch (Throwable t) \{
throw new RuntimeException("An error occured while executing "
+ "doInBackground()", t);
\}
...
\}
\}

%SELECT (exception, root_cause) AS PAIR, COUNT (*) AS N FROM stacktracesummary where 
%EXCEPTION_EXTENDS='RUNTIME' AND ROOT_EXTENDS='ERROR' GROUP BY PAIR ORDER BY N DESC;

(java.lang.RuntimeException,java.lang.OutOfMemoryError) \- 30
(java.lang.RuntimeException,java.lang.NoClassDefFoundError) \- 6
(java.lang.RuntimeException,java.lang.UnsatisfiedLinkError) \- 5
(android.view.InflateException,java.lang.OutOfMemoryError) \- 4
(java.lang.RuntimeException,java.lang.StackOverflowError) \- 4

TO INVESTIGATE: WHICH WERE THE LIBS AND THE APP? AND ITS REPUTATION?

\subsubsection{Checked wrapping Runtime Exception}

APP	9
LIB	9
ANDROID	4

(com.octo.android.robospice.persistence.exception.CacheSavingException,java.lang.ClassCastException)	6
(java.lang.reflect.InvocationTargetException,java.lang.NullPointerException)	3
(java.lang.reflect.InvocationTargetException,android.content.res.ResourcesdNotFoundException)	2
(java.sql.SQLException,android.database.sqlite.SQLiteDiskIOException)	2
(java.lang.Exception,at.andiwand.odf2html.odf.IllegalMimeTypeException)	2

The most frequent exception wrapping were were performed inside RoboSpice is an android library that supports the creation of asynchronous long running tasks.
This library has 1.682 stars and 367 forks.

Most of such wrappings however were caused by the reflection Java library, used by Android, Application and library classes.
relaction methods convert every exception thata happens when invoking a reflected method into an INVOCATION TARGET EXCEPTION
this conversion is performed inside native code.

\subsubsection{Checked Exception wrapping Error}

We could observe that in 7 projects this  wrapping happend 8 times in total.
Inspecting the stack traces we could observe that 7 out of 8 wrappings were caused
by the reflection java library on which java.lang.OutOfMemoryError
java.lang.NoSuchMethodError were wrapped into a java.lang.reflect.InvocationTargetException.

ANDROID	5
APP	3

(java.lang.reflect.InvocationTargetException,java.lang.OutOfMemoryError)	5
(java.lang.Exception,java.lang.OutOfMemoryError)	1
(java.lang.reflect.InvocationTargetException,java.lang.NoSuchMethodError)	1
(java.lang.reflect.InvocationTargetException,java.lang.UnsatisfiedLinkError)	1

All wrappings except 1 were caused by the reflection library.

at java.lang.reflect.Constructor.constructNative(Native Method)
NATIVE IMPLEMENTATION, WE DO NOT HAVE ACCESS TO THE SOURCE CODE… IT IS IN C…
CHECKED EXCEPTIONS CAN BE THROWN BY NATIVE METHODS…. ANY CHANING IS PERFORMED?

In one of the cases, however, in project OpenDocument.droid a java.lang.Exception was 
wrapping a java.lang.OutOfMemoryError. General handlers are considered as bad practice, 
but wrapphing any exception on a general Exception make the matters worse.

\subsubsection{Error wrapping Checked Exception}

%SELECT (exception, root_cause) AS PAIR, COUNT (*) AS N FROM stacktracesummary where 
%EXCEPTION_EXTENDS='ERROR' AND ROOT_EXTENDS='CHECKED' GROUP BY PAIR ORDER BY N DESC;

libcore	\- 3
 JAVA ENV \- 	4

(java.lang.NoClassDefFoundError,java.lang.ClassNotFoundException);4
(java.lang.AssertionError,javax.crypto.ShortBufferException);3

All ClassDefFound wrappings were caused by java.lang.Class methods such as 
at java.lang.Class.getDeclaredFields(Native Method)
at java.lang.Class.getDeclaredConstructors(Native Method). They are all NATIVE METHODS.

\subsubsection{Error wrapping Runtime}

APP	11
ANDROID	1
LIB	1

@@

Since there is no way of enforcing Java exception type conventions during program development,
the exceptions regardless its types can se used interchangeably. This interchangeability makes the
behaviour of the exception handling code more complex and less reliable. For
instance, considering a situation were an instance of OutOfMemoryError (a situation that usually cannot
 be handled) was wrapped in a checked exception which tells the caller that such exception should
handled.

@@

The exception (mis-)use patterns that emerged  in this study sheds light on 
problems that may be reducing the robustness of current Java programs in the face of
exceptional conditions.

%\noindent\emph{The undocumented runtime exceptions} 
\subsection{The undocumented runtime exceptions thrown by thrid party code} 
Documenting runtime exceptions is a tedious and error prone task, to help developers
mitigate this problem tools should be developed to automate the documentation of runtime exceptions
scraping from library code, few solutions in this directions have been proposed so far ~\cite{van2005combining}.

%\noindent\emph{The odd wrappings}  
\subsection{Exception Wrappings: Language design purpose x real use}

The exception wrappings were inicially proposed by Java [ref-specification] to ...

But the types of wrappings found in this study pointing to the fact that some wrappings may have not 
been used according to its intial purpose. It can sometimes be used to bypass the  
language restrictions imposed by checked exceptions.

When (mis)applied, the exception wrapping can lead to what we call
an \emph{exception handling confusion problem} which can lead the program to
an unpredictable state in the presence of exceptions. To illustrate this problem, we can use
 one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the JVM which make impossible the program to recover. Further investigation is needed on finding 
ways to help developers in deadling with such problem, either preventing odd wrappings or enabling 
the developer to better deal with them.

%\noindent\emph{The high number of null pointer exceptions}

\subsection{The high number of null pointer exceptions}  
Last but not least, the high number of null pointer exceptions  lead us to think on other 
pertinent research direction: (i) lightweight intra-method nullpointer analysissuch as the @Nullable;
(ii) intermethod nullpointer analysis tools such as the one proposed by [DANIELE-RELATED-WORK];
or (iii) whether language designs which avoid null pointers, such 
as Monads ~\cite{Walde95} (i.e., used in functional languages for values that may not be available 
or computations that may fail) could improve the robustness of Java programs. 
@NULABLE ANNOTATIONS.

%\noindent\emph{Dependency injection - Inversion of Control -  callback and Exceptions }  
\subsection{Dependency injection - Inversion of Control -  callback and Exceptions}

%\noindent\emph{How can we benefit from such cross-project crash analysis?}  
\subsection{The benefits of a cross-project crash analysis}

When family of systems make use of a common core, or framework such systems can
bennefit from a cross project crash analysis. The bug issues reported for a group of systems
can identify vulnerabilities which were more difficult to identify in a single project analysis.

The test case explosion problem... A test suite is always incomplete,
one application could benefit from the usage scenarios on which such framework is applied.
Framework testing -  what are the main challenges wor framework testing?
A crossproject crash analysis based on repository mining techniques can 
identify vulnerabilitues in pieces of code that are shared.

Using exception miner tool we could identify the methods reponsible for most of the crashes - 
such methods can be a target for manual inspections improvements, or better documentation
(because sometimes an exception can be thrown because the user is not invoking the method
in the proper way

During such analysis we could also find the undocumented runtime exceptions 
signaled by libraries.

It allows the developers to benefit from the faults happening on different systems.

@@

\begin{table}
\centering
\begin{tabular}{llll}
    \hline
    \bfseries{Project} & \bfseries{merging imeStars} & \bfseries{Watchers} & \bfseries{Forks} \\
    \hline
subsampling-scale-image-view  &  -  &  -    & - \\
Android-PullToRefresh &  -  &  -    & - \\
HackerNews &  -  &  -    & - \\
PathfinderOpenReference  &  -  &  -    & - \\
AndroidStaggeredGrid  &  -  &  -    & - \\
HtmlSpanner  &  -  &  -    & - \\
screen-notifications  &  -  &  -    & - \\
appdatapreferences-android  &  -  &  -    & - \\
Android-Universal-Image-Loader  &  -  &  -    & - \\
Android-SDK  &  -  &  -    & - \\
WebCachedImageView  &  -  &  -    & - \\
Android-SkyDrive-Browser  &  -  &  -    & - \\
android-pdfview  &  -  &  -    & - \\
hn-android  &  -  &  -    & - \\
Android-BitmapCache  &  -  &  -    & - \\
cropper  &  -  &  -    & - \\
WordPress-Android  &  -  &  -    & - \\
changeloglib  &  -  &  -    & - \\
cgeo  &  -  &  -    & - \\
chromium-webview-UNDERLINE  &  -  &  -    & - \\
AndroidSlidingUpPanel  &  -  &  -    & - \\
DsaTab  &  -  &  -    & - \\
box-android-sdk-v2  &  -  &  -    & - \\
chromeview  &  -  &  -    & - \\
andlytics  &  -  &  -    & - \\
ContractionTimer  &  -  &  -    & - \\
android  &  -  &  -    & - \\
okhttp  &  -  &  -    & - \\
ActionBarSherlock  &  -  &  -    & - \\
mobile-android  &  -  &  -    & - \\
android-timepicker  &  -  &  -    & - \\
QuiltViewLibrary  &  -  &  -    & - \\
congress-android  &  -  &  -    & - \\
c-geo-opensource  &  -  &  -    & - \\
TweetLanes  &  -  &  -    & - \\
PageTurner  &  -  &  -    & - \\
SWADroid  &  -  &  -    & - \\
picasso  &  -  &  -    & - \\
uservoice-android-sdk  &  -  &  -    & - \\
\hline
  \end{tabular}
\caption{Projects containing wrappings Error-Runtime}
\label{tab:wrappings3}
\end{table}

The methods performing such wrappings were the following:

%source code of ActivityThread -  https://github.com/android/platform_frameworks_base/blob/master/core/java/android/app/ActivityThread.java

\begin{table}
\begin{tabular}{l}
    \hline
    \bfseries{root signaler -  occurences} \\
      \hline
android.app.ActivityThread.handleBindApplication – 2 \\
android.app.ActivityThread.performLaunchActivity – 13 \\
android.os.AsyncTaskd3.done – 25	 \\
android.support.v4.content.ModernAsyncTaskd3.done - 2 \\
android.view.LayoutInflater.createView - 7 \\
android.view.Viewd1.onClick - 1 \\
com.squareup.picasso.Requestd1.run - 1 \\
me.openphoto.android.app.util.concurrent.AsyncTaskExd3.done - 1 \\
org.asynctasktex.AsyncTaskExd3.done – 3 \\
org.chromium.androidwebview.AwBrowserProcess.loadLibrary -1 \\
org.gradle.groovy.scripts.internal.DefaultScriptRunnerFactorydScriptRunnerImpl.run - 1 \\
\hline
  \end{tabular}
\caption{Top 10 most popular exceptions on issues in Java projects}
\label{tab:toptenjava}
\end{table}

@@
